---
title: RGL + Rmarkdown Tricks
author: Susan Vanderplas
date: '2019-05-22'
categories:
  - R
tags:
  - note-to-self
  - useful-code
image:
  caption: ''
  focal_point: ''
---


Basic setup - requires `x3ptools` and `rgl` packages
```{r setup}
library(x3ptools)
library(rgl)
logo <- read_x3p(system.file(package = "x3ptools", "csafe-logo.x3p"))

rgl::setupKnitr()
```

## Changing viewing angle

You can interactively get the viewing angle by using `rgl::par3d()$userMatrix`

For example, if I use `image_x3p(logo)` to show the CSAFE logo, and then I rotate the object a bit, I can use

```{r, eval = F}
# Not run...
um <- rgl::par3d()$userMatrix
dput(um)
```

to save the rotation matrix for reproducibility. I can then set it using `view3d(userMatrix = um)`


## RGL and Rmarkdown (static)

- chunk option: rgl = TRUE
- chunk option: dev = png

Allows you to use RGL and generate a static image automatically. The chunk header looks like this^[
Note: Get verbatim chunks using [this post](https://yihui.name/en/2017/11/knitr-verbatim-code-chunk/)]:
````
```{r, rgl = TRUE, dev = 'png'}`r ''`
## Chunk code goes here
```
````

```{r, eval = T, rgl = TRUE, dev = 'png', fig.width = 8, fig.height = 4, out.width = "100%"}
um <- structure(c(0.990302324295044, -0.108039885759354, -0.0873337239027023, 
                  0, 0.119859784841537, 0.982302069664001, 0.143926605582237, 0, 
                  0.0702383369207382, -0.152998745441437, 0.985726416110992, 0, 
                  0, 0, 0, 1), .Dim = c(4L, 4L)) # from dput()
image_x3p(logo)
rgl::view3d(userMatrix = um)

scene <- rgl::scene3d() # save the scene to an R object
```


[Stackoverflow post](https://stackoverflow.com/questions/53892177/how-to-embed-an-rgl-snapshot-in-rmarkdown-package-vignette)

## RGL and Rmarkdown (dynamic)

To get RGL embedded in RMarkdown (warning, this will make your HTML files extremely large...), use `rgl::rglwidget()` 
```{r plot-dynamic, fig.width = 8, fig.height = 4, out.width = "100%"}
# How to plot:
scene %>%
  rglwidget()
```

- [RGL/WebGL vignette](https://cran.r-project.org/web/packages/rgl/vignettes/WebGL.html)
