---
title: Exploring GAMs
author: Susan Vanderplas
date: '2020-11-22'
slug: []
categories:
  - R
tags:
  - R
  - new-techniques
subtitle: ''
summary: ''
authors: []
---


```{r setup, message = F, warning = F}
library(tidyverse)
library(skimr)
library(mgcv)
```

I'm going to start out playing with the nasa data from `ggpcp` - it's spatial in nature and should be well handled by a nonparametric model.

```{r}
data("nasa", package = "ggpcp")
```

A basic model might look at the following relationships:

```{r}
library(mgcv)

ggplot(nasa, aes(x = elevation, y = temperature - 273)) + 
  geom_point() + 
  geom_smooth(method = "gam", formula = y ~ s(x), se = T) + 
  ylab("Temperature (deg C)") + xlab("Elevation (m)")

ggplot(nasa, aes(x = elevation, y = temperature - 273, color = month)) + 
  geom_jitter() + 
  geom_smooth(method = mgcv::gam, formula = y ~ s(x), se = F) + 
  ylab("Temperature (deg C)") + xlab("Elevation (m)")

ggplot(nasa, aes(x = lat, y = temperature - 273)) + 
  geom_jitter() + 
  geom_smooth(method = mgcv::gam, formula = y ~ s(x), se = F) + 
  ylab("Temperature (deg C)") + xlab("Latitude")

ggplot(nasa, aes(x = long, y = temperature - 273)) + 
  geom_jitter() + 
  geom_smooth(method = mgcv::gam, formula = y ~ s(x), se = F) + 
  ylab("Temperature (deg C)") + xlab("Longitude")
```

So let's fit a basic model - smooth elevation by month, and then add in an interaction tensor between latitude and longitude.

```{r}
basic_model <- gam(data = nasa, temperature - 273 ~ s(elevation, by = month) + te(lat, long), method = "REML")
```

We can check diagnostics with `gam.check`:

```{r}
gam.check(basic_model)
```

For the most part it looks ok, but most of our k-indexes are <1 and some of the EDF are >8 when k' is 9. So lets see what changes with the gam_check output if we add k=15. No clue what will happen, but that's what makes this interesting.

```{r}
basic_model2 <- gam(data = nasa, temperature - 273 ~ s(elevation, by = month) + te(lat, long, k = 15, bs = "cs"), method = "REML")
```

```{r}
gam.check(basic_model2)
```

And we can also see what happens when we increase k on the elevation smoother. 

```{r}
basic_model3 <- gam(data = nasa, temperature - 273 ~ s(elevation, by = month, k=10) + te(lat, long), method = "REML")
```

```{r}
gam.check(basic_model3)
```
